---
title: "Week 4 Presentation"
author: "Emma Li"
format: html
editor: visual
---

## Program - Chapter 25, 26 and 27

-   Functions
    -   Vector functions
    -   Data frame functions
    -   Plot functions
-   Iteration
    -   Modify multiple columns with dplyr::across()
    -   Read multiple files with purrr::map()
    -   Save multiple outputs
-   Important base R functions
    -   Selecting element(s) with \[\], \$, \[\[\]\]
    -   Apply functions, base R equivalent of map functions
    -   For loops
    -   Base R plots

## Effects of Intensive Glucose Lowering in Type 2 Diabetes

-   New England Journal of Medicine
-   Published June 12, 2008

### Statistical Analysis Methods

-   Power and Sample Size
    -   Designed with 89% likelihood to detect whether intensive therapy will lower rates of serious cardiovascular events by 15% assuming standard therapy has annual event rate of 2.9%
-   Data Analysis
    -   Checked that 2 groups were similar at the start of the study using chi-square tests and two-sample t-tests
    -   Compared how long patients in each group experienced primary outcomes using Cox proportional-hazards regression models
    -   Calculated risk of primary and secondary outcomes in both groups with 95% confidence of the estimates
    -   Used Kaplan-Meier estimates to find the percentage of patients experiencing events over time in each group

### Main Findings

-   No significant reduction in primary outcomes
-   More mortality in intensive treatment group

## ACCORD Data Exercise - Numerically and Visualing Describing the Percentages of Different Groups

-   Question: What's the percentage of females and of each ethnicity group in `accord_key`? Describe the results numerically and visually.
    -   Created a function to calculate percentage
    -   Applied it to all groups (female, White, Black, Hispanic, Other)
    -   Graphed the percentages

```{r, echo=FALSE, results='hide', warning=FALSE, message=FALSE}
rm(list=ls())
knitr::opts_chunk$set(echo = TRUE)
library(dplyr)
library(ggplot2)
library(gridExtra)
library(gtsummary)
library(patchwork)
library(readr)
library(readxl)
library(sjPlot)
library(tidyverse)

datasets_folder <- "/Users/emmali/Desktop/Emma's Stuff/Summer Work/2024/UCLA Internship Work/ACCORD_exercises/datasets/"

accord_key <- read_csv(paste(datasets_folder, "accord_key.csv", sep = "")) |>
  print(width = Inf)
bloodpressure <- 
  read_csv(paste(datasets_folder, "bloodpressure.csv", sep = "")) |>
  print(width = Inf)
hba1c <- read_csv(paste(datasets_folder, "hba1c.csv", sep = "")) |>
  print(width = Inf)

accord_key <- accord_key |>
  rename(gender = female)
accord_key$gender <- ifelse(accord_key$gender == 1, "Female", "Male")
raceclass_levels <- c("White", "Black", "Hispanic", "Other")
accord_key$raceclass <- accord_key$raceclass |>
  factor(levels = raceclass_levels)
```

```{r, warning=FALSE, message=FALSE}
num_of_patients <- accord_key |>
  nrow()
calculate_percent <- function(data, column, desired_value) {
  percent <- data |> 
    filter({{column}} == desired_value) |> 
    nrow()
  percent / num_of_patients * 100
}
percent_female <- calculate_percent(accord_key, gender, "Female")
percent_white <- calculate_percent(accord_key, raceclass, "White")
percent_black <- calculate_percent(accord_key, raceclass, "Black")
percent_hispanic <- calculate_percent(accord_key, raceclass, "Hispanic")
percent_other <- calculate_percent(accord_key, raceclass, "Other")
```

-   percentage of female patients: `round(percent_female, digits = 2)`
-   percentage of White patients: `round(percent_white, digits = 2)`
-   percentage of Black patients: `round(percent_black, digits = 2)`
-   percentage of Hispanic patients: `round(percent_hispanic, digits = 2)`
-   percentage of other patients: `round(percent_other, digits = 2)`

```{r, warning=FALSE, message=FALSE}
percentages = data.frame(groups = c("female", "white", "black", "hispanic", "other"), percents = c(percent_female, percent_white, percent_black, percent_hispanic, percent_other))
ggplot(percentages, aes(x = groups, y = percents)) +
  geom_col(stat = "identity") +
  labs(title = "Percentages of Females and Ethnic Groups",
       subtitle = "More than 50% of patients are White, less than 7% of patients are Hispanic",
       x = "Groups",
       y = "Percentages")
```
