---
title: "week_7_presentation"
format: html
editor: visual
---

## Mastering Shiny

### Chapter 9

-   Uploads
    -   fileInput() can be added to ui to upload a file, requires id and label
    -   In server, it returns data frame with name, size, type and datapath
    -   When uploading a dataset:
        -   Add req(input\$upload) in server
        -   (optional) add accept argument to fileInput() to limit what kinds of files can be uploaded
-   Downloads
    -   downloadButton() or downloadLink() can be added to ui, requires id
    -   No render functions, use downloadHandler() what file to download and where
    -   Download reports using parameterised RMarkdown document

### Chapter 10

-   Dynamic UI
    -   Update functions to update inputs
        -   Every input control has a update function
        -   Ex. textInput() and updateTextInput()
        -   Hierarchical select boxes
        -   Freezing reactive inputs
            -   Hierarchical selection can create invalid inputs which lead to flickering outputs
            -   freezeReactiveInput() stops reactive inputs from updating when its not necessary and prevents errors
        -   Circular references
            -   Be mindful to not create infinite loops when updating values
        -   Inter-related inputs
    -   Dynamic visibility
        -   Use tabset panel with hidden tabs and updateTabsetPanel() to switch tabs to show and hide different parts of the UI
        -   Conditional UI
            -   Depending on what the user selects, display different options
        -   Wizard interface
            -   A kind of interface that makes it easier to collect information by spreading it across many pages
    -   Creating UI with code
        -   uiOutput() inserts placeholder in ui
        -   renderUI() fills in the placeholder from server
        -   Multiple controls with purrr functions map() and reduce()

### Chapter 11

-   Bookmarking
    -   Shiny doesn't expose the current state of an app in the URL
    -   How to make an app bookmarkable:
        -   Add bookmarkButton() to ui
        -   Turn ui into a function
        -   Add enableBookmarking = "url" to shinyApp()
    -   You can also automatically update the URL
    -   Adding enableBookmarking = "server" to shinyApp() saves state to .rds file

## Predictive Modeling

### Logistic Regression

Predict food insecurity status using household’s socio-economical status

```{r}
library(readr)
library(gtsummary)
library(knitr)
library(pROC)
library(tidymodels)
```

Loading the data

```{r}
data_clean <- read_rds("../Week 6-7/fss21.rds") |>
  print()
```

Using logistic regression to model the probability of a household being in low food security

```{r}
logit_model <- glm(
  # All predictors except HRFS12M1 and HHSUPWGT
  HRFS12M1_binary ~ . - HRFS12M1 - HHSUPWGT,
  data = data_clean, 
  family = "binomial"
  )
logit_model
```

Summary of the model with gtsummary

```{r}
logit_model |>
  tbl_regression() |>
  bold_labels() |>
  bold_p(t = 0.05)
```

Importance of each predictor evaluated using the ANOVA test

```{r}
anova(logit_model, test = "Chisq")
```

### ROC Curve

Using different thresholds when using the model for prediction

```{r}
pred_prob <- predict(logit_model, type = "response")

# Set the threshold 0.5
threshold <- 0.5

predicted_class <- ifelse(pred_prob > threshold, 1, 0) |> as.factor()
actual_class <- data_clean$HRFS12M1_binary

ctb_50 <- table(Predicted = predicted_class, Actual = actual_class)

# Set the threshold 0.1
threshold <- 0.1

predicted_class <- ifelse(pred_prob > threshold, 1, 0) |> as.factor()
actual_class <- data_clean$HRFS12M1_binary

ctb_10 <- table(Predicted = predicted_class, Actual = actual_class)

list(threshold50 = ctb_50, threshold10 = ctb_10)
```

Calculating the metrics of using 0.5 verses 0.1 as the threshold

```{r}
# Calculate metrics under different thresholds
calc_metrics <- function(ct) {
  c(Accuracy = (ct[1, 1] + ct[2, 2]) / sum(ct),
    Sensitivity = ct[2, 2] / sum(ct[, 2]),
    Specificity = ct[1, 1] / sum(ct[, 1]),
    Precision = ct[2, 2] / sum(ct[2, ]))
}

metrics50 <- calc_metrics(ctb_50)
metrics10 <- calc_metrics(ctb_10)

data.frame(threshold50 = metrics50, threshold10 = metrics10) |>
  t() |>
  round(2) |>
  kable()
```

From the metrics: A threshold of 0.5 has a high accuracy rate (0.90) but misses many cases that it is designed to capture (in this case, households with low food security, this model only captures 2% of them). A threshold of 0.1 has a lower accuracy rate (0.74) but captures a lot more cases that it should (71% of households with low food security).

Graphing a ROC curve and calculating the AUC

```{r}
data_clean <- data_clean |>
  mutate(prob = predict(logit_model, type = "response"))

roc_data <- roc(data_clean$HRFS12M1_binary, data_clean$prob)

ggroc(roc_data, legacy.axes = TRUE) +
  labs(title = "ROC Curve for Logistic Regression Model",
       x = "1 - Specificity",
       y = "Sensitivity") +
  theme_minimal() +
  annotate("text", x = 0.5, y = 0.5,
           label = paste("AUC =", round(auc(roc_data), 3)))
```

AUC is 0.794 which is pretty good. Ideally, we want the AUC to be as close to 1 as possible.

Use Mean Square Error and Cross Validation to determine how good the model is

### Logistic Regression with Enet (elastic-net) Regularization

Split the data

```{r}
# For reproducibility
set.seed(2024)

data_split <- data_clean |>
  initial_split(
  # stratify by HRFS12M1_binary
  strata = "HRFS12M1_binary", 
  prop = 0.75
  )
data_split

data_other <- training(data_split)
dim(data_other)

data_test <- testing(data_split)
dim(data_test)
```

Recipe

```{r}
recipe <- recipe(
    HRFS12M1_binary ~ .,
    data = data_other
  ) |>
  # remove the weights and original HRFS12M1
  step_rm(HHSUPWGT, HRFS12M1) |>
  # create dummy variables for categorical predictors
  step_dummy(all_nominal_predictors()) |>
  # zero-variance filter
  step_zv(all_numeric_predictors()) |> 
  # center and scale numeric data
  step_normalize(all_numeric_predictors()) |>
  # estimate the means and standard deviations
  print()
```

Model

```{r}
logit_mod <- logistic_reg(
    penalty = tune(), # \lambda
    mixture = tune()  # \alpha
  ) |> 
  set_engine("glmnet", standardize = FALSE) |>
  print()
```

Workflow

```{r}
train_weight <- round(data_other$HHSUPWGT / 1000, 0)
train_weight <- ifelse(train_weight == 0, 1, train_weight)

logit_wf <- workflow() |>
  #add_case_weights(train_weight) |>
  add_recipe(recipe) |>
  add_model(logit_mod) |>
  print()
```

Tuning Grid

```{r}
param_grid <- grid_regular(
  penalty(range = c(-3, 3)), # \lambda
  mixture(), # \alpha
  levels = c(1000, 5)
  ) |>
  print()
```

Cross Validation

Set partitions

```{r}
(folds <- vfold_cv(data_other, v = 5))
```

Fit cross validation

```{r}
(logit_fit <- logit_wf |>
  tune_grid(
    resamples = folds,
    grid = param_grid,
    metrics = metric_set(roc_auc, accuracy)
    )) |>
  system.time()
```

Visualize CV result

```{r}
logit_fit |>
  # aggregate metrics from K folds
  collect_metrics() |>
  print(width = Inf) |>
  filter(.metric == "roc_auc") |>
  ggplot(mapping = aes(
    x = penalty, 
    y = mean, 
    color = factor(mixture)
    )) +
  geom_point() +
  labs(x = "Penalty", y = "CV AUC") +
  scale_x_log10()
```

5 best models

```{r}
logit_fit |>
  show_best(metric = "roc_auc")
```

Best model

```{r}
best_logit <- logit_fit |>
  select_best(metric = "roc_auc")
best_logit
```

Final Model

```{r}
# Final workflow
final_wf <- logit_wf |>
  finalize_workflow(best_logit)
final_wf
```

```{r}
# Fit the whole training set, then predict the test cases
final_fit <- final_wf |>
  last_fit(data_split)
final_fit
```

```{r}
# Test metrics
final_fit |> 
  collect_metrics()
```
